# PSCI 1800 Midterm

**Student ID: **

Opens: 12:00pm on Wednesday March 4th

Due: 1:00pm on Wednesday March 4th.

## Instructions:

  -	This is an in-class open-note test. You have 1 hour to complete the exam.
  
  -	I do not accept late work. Exams handed in after 60 minutes will not be graded.

  -	The test is open notes. You can use any course material you wish, including this textbook, your problem sets, the problem set answer keys, and your own notes. You cannot google things, and all use of AI is prohibited. 

  - As with the problem sets, you will be graded on your  **comprehension of the material in this specific course**.

  -	You may not use Chat-GPT or any other AI tool to answer the questions. Violating this policy will result in a score of 0 for the midterm and an immediate referral to the Center for Community Standards \& Accountability to decide further appropriate disciplinary action.

  -	Your exam submission will be identical to how you've submitted problem sets. You will submit a .rmd file and a knitted html.

  - The exam will be graded anonymously. Please put your student number on the exam only.


Good luck!

_____

We are going to work with panel data from the Cooperative Election Study (CES). The CES is a large national survey that asks Americans about their political opinions and behavior. In this dataset each respondent was surveyed twice: once in 2010 and once in 2014. 

Along with some stable basic demographic information, we have a measure of respondents' fiscal policy preferences `deficit.fix` in each interview year. This is a 100 point scale asking how the federal budget deficit should be reduced, where 0 means "all from tax increases" (the liberal option) and 100 means "all from spending cuts" (the conservative option).

There is no missing data in this dataset.

You can load the data in here:

```{r, class.source="bg-success"}
dat <- rio::import("https://github.com/marctrussler/IDS-Data/raw/refs/heads/main/PSCI1800Midterm2026DataLong.Rds", trust=T)
```


**1. How many respondents are in these data? How many rows are in these data? Based on that information, and by looking at the dataset, what is the unit of analysis of these data?**


**2. Un-comment and edit the code below to reshape the data so that the unit of analysis is the individual respondent. Two new variables will be created in the process. (You do not need to edit the `names_prefix` option, which I've added so that we get the same sensible variable names to work with going forward.)**

```{r}
library(tidyr)
#dat <- pivot_wider(dat,
#                   ??????,
#                   ??????,
#                   names_prefix = "budget_deficit_scale.")
```

***If you are not able to successfully complete this step, use this code to load in a re-shaped version of the data so you can continue with the exam***

```{r}
dat <-  rio::import("https://github.com/marctrussler/IDS-Data/raw/refs/heads/main/PSCI1800Midterm2026DataWide.Rds", trust=T)
```


**3. What is the correlation between people's birth year and `deficit.fix.2010`? What is the correlation between people's birth year and `deficit.fix.2014`? Create a new variable `deficit.change` which finds the difference between a person's opinion on this question in 2010 and in 2014. This variable should be calculated such that positive values indicate that someone has moved in a conservative direction (i.e. towards preferring spending reductions).  What is the correlation between people's birth year and this new variable? What do you conclude from these three correlations?**


**4. What proportion of men born before or during 1975 moved in a conservative direction (towards wanting spending cuts) from 2010 to 2014? What proportion of men born after 1975 moved in a conservative direction?  What does this tell you? *Hint: What does the mean of a boolean variable represent?*** 


**5. Edit the code below to answer this question: *For* each unique county in the dataset, find the maximum value for `deficit.change` and the minimum value for `deficit.change`. Which county has the biggest difference between the individual with the maximum and minimum change? *Hint: an easy mistake to make here is to not identify all the unique counties in the data.***

```{r}
#county <- ?????
#max.deficit.change <- rep(NA,???????)
#min.deficit.change <- rep(NA,???????)

### 
#??????
###

#deficit.delta <- max.deficit.change - min.deficit.change

```

**6. Below I calculate the mean and standard deviation of `deficit.fix.2010`. What is the standard error of this sample mean? In words, what does that mean?**

```{r}
mean(dat$deficit.fix.2010)
sd(dat$deficit.fix.2010)
```

